Our Future Nuclear Data Needs

### 2018-12-16 nuclear_data_needs/nuclear_data_needs.tex
Finished through section 2.1.  Dumping rest of text in

\usepackage[american]{babel}
% \usepackage{amsmath}
% \usepackage[version=3]{mhchem} 
% % \usepackage{fixltx2e}
% % \usepackage{refcount}
% % \usepackage{siunitx}
% % \usepackage{lastpage}
% % \usepackage{textcomp}
% \usepackage{mathtools}
% 
% \usepackage{xfrac}
\usepackage{lmodern}
\usepackage[hidelinks]{hyperref}
% % \usepackage{cool}
% % \usepackage{cancel}
% % \usepackage{microtype}
% \usepackage{listings}
% % \usepackage{mcode}
\usepackage [autostyle, english = american]{csquotes}
% % \usepackage{longtable}
% % \usepackage{subcaption}
% \usepackage{booktabs,siunitx}
% \usepackage{gensymb}
% \usepackage[normalem]{ulem}

@@

@@

%Abstract
\begin{abstract}
A well-established knowledge of nuclear phenomena including fission, reaction cross sections, and structure/decay properties, is critical for applications ranging from the design of new reactors, to non-proliferation, to the production of radioisotopes for the diagnosis and treatment of illness.  However, the lack of a well-quantified, predictive theoretical capability means that most nuclear observables must be measured directly in order to “calibrate” empirical models, which, in turn, provide the data needed for these applications.  In many cases there is either a lack of the data needed to guide the models, or the results of the different measurements are discrepant.  This has led to the development of evaluation methodologies to provide recommended values and uncertainties.  In this article we present the nuclear data evaluation process and the international community that carries it out.  We then discuss new measurements and improved theory and/or modeling needed to address future challenges in applied nuclear science. 
\end{abstract}

@@

The nuclear structure compilation database is called the Experimental Unevaluated Nuclear Data List (XUNDL) \cite{XUN}.   The nuclear data in XUNDL is organized by nuclide, meaning that a single publication could lead to the production of more than one XUNDL dataset.  Each XUNDL dataset is treated as a “stand-alone” work and is not required to agree with existing, evaluated nuclear data for the nuclide it concerns.  However, during XUNDL compilation basic consistency checks are performed and any internal inconsistencies are pointed out to the author to allow them be reconciled in the database.  XUNDL is used by both nuclear structure evaluators to aid in their work and by nuclear structure researchers as a quick way to get access to data from current publications in order to guide their own research.

@@

While ENSDF and ENDF contain a vast quantity of nuclear data, there are significant amounts of nuclear data that are either not contained in them because they “fall between” nuclear structure and nuclear reactions, or they are present in a format that is not well-suited to their use in applications.  Examples of this include the recently updated Atlas of Neutron Resonances \cite{Mug18} which contains neutron capture resonance widths, centroids and cross sections and the Evaluated Gamma Activation File of (EGAF) capture gamma-rays \cite{Fir15} and the Atlas of Inelastic Scattering of Reactor Fast Neutron \cite{Hur18}.  Another very important nuclear data resource for applications is the Reference Input Parameter Library (RIPL) \cite{Capote2009} which contains both discrete (distilled from ENSDF) and continuous (Nuclear Level Densities and Gamma-strength function) nuclear structure data needed for accurate reaction modeling.

@@

The final step in the Pipeline is validation.  In this step, we use user codes and user-defined problems to validate that the data performs as expected.  The class of user problems we most rely on are so-called “integral benchmark” data that depend on many different types of data including cross sections and emitted particle spectra and are also extremely precisely known.  The canonical example of such a benchmark is a critical assembly, where the ratio of neutron production to neuron loss (k­eff) is known to several significant figures.  However, there are many other kinds of integral benchmarks used to test neutron interaction data alone, including reaction rate measurements and 14 MeV and fission spectrum transmission experiments, and many more for other forms of nuclear data.

There are multiple ways to interact with the nuclear data evaluation process along the way, including through web interfaces hosted by the National Nuclear Data Center \cite{NNDC} and the IAEA Nuclear Data Section \cite{IAEA-NDS}, a smartphone based application \cite{IAEA-APP} and publication in peer-reviewed journals including Nuclear Data Sheets \cite{NDS} and Atomic Data and Nuclear Data Tables \cite{ATNDT}.

@@

Much of the work of nuclear data evaluation takes place as a part of international collaborations covered under the Organization of Economic Cooperation and Development’s Nuclear Energy Agency (NEA) and the International Atomic Energy Agency (IAEA).  The Nuclear Energy Agency (NEA) addresses the nuclear technology interests of its 33 member states and, in the area of nuclear data, the Working Party on Evaluation Coordination (WPEC) is the main forum for collaborative effort between the nuclear data library projects from the NEA countries, namely ENDF/B (United States), JENDL (Japan), JEFF (NEA), TENDL (Europe), BROND (Russia) as well as the non-OECD file project CENDL (China).

A powerful instrument of WPEC to drive progress in nuclear data is the so-called Subgroup (SG): members of WPEC identify a common area in nuclear data that requires improvement, and if enough support from the various data library projects is present, a Subgroup is formed.  Subgroups typically operate on a 3-5 year time frame. Recent successful examples of WPEC subgroups, which specific relevance to the US, are large-scale horizontal efforts like CIELO (SG40), on worldwide nuclear data evaluation for the most important fission energy related materials. The CIELO initiative has for example led to new ENDF/B-VIII.0 evaluations for 235,238U, 56Fe and 16O among others, and the Generalized Nuclear Database Structure (GNDS) format (SG38) \cite{SG38}, or more specific efforts like covariance adjustment for improvement of nuclear data files (SG39) \cite{SG39}.

WPEC also hosts long-term Expert Groups, and current groups are working on the development of the Generalized Nuclear Database Structure (GNDS) format, to provide a data library interface between nuclear physics and applications more modern than the ENDF-6 format, and the High-Priority Request List which assembles the most important nuclear data requests from applications in a unified format, to stimulate experimentalists and evaluators to provide these data. A full list of past and current subgroups of WPEC is available at Ref. \cite{WPEC}.
 
The International Atomic Energy Agency (IAEA) covers the interests of its 170 member states. The main task of the IAEA Nuclear Data Section is to provide fundamental nuclear data bases for basic and applied use, with data originating from experiments and theoretical simulations and covering both nuclear structure and nuclear reaction data.  An important collaboration coordinated by the IAEA is the Nuclear Reaction Data Center Network, which is responsible for keeping the EXFOR database of experimental nuclear reaction data up to date. NNDC is responsible for the US input to the database.

In addition, the IAEA organizes Coordinated Research Projects (CRP) and technical meetings as instruments to align international nuclear data efforts towards the production of validated databases ready for applied use.  Examples of recent and current CRP's include a 2018 venture centered on nuclear data for primary radiation damage has been completed, an ongoing effort to improve nuclear model parameters for fission reaction calculations by modern nuclear model codes such as EMPIRE \cite{Her07}, CCONE \cite{Iwa16}, COH3 \cite{Kaw10} and TALYS \cite{Kon08} is ongoing, and an effort to create a first-ever evaluated database of radiative strength.  In 2019, a CRP on fission yields will start which aims to produce updated fission yield libraries for the major actinides to respond to requests from reactor technology, safeguards and non-proliferation.

@@

Accurate, physics-based modeling of neutron-induced reactions, particularly on the “Big 3” (235,238U and 239Pu) are key to virtually all nuclear science and technology applications.  While a great deal of effort has been made to improve (n,f) data, a fundamental understanding of this most important of reaction mechanisms remain elusive.  Furthermore, since fission leads to the production of neutrons with energies up to several MeV, elastic and inelastic scattering also play a critical role.  Unfortunately, efforts to improve neutron scattering data has tended to “take a back seat” to fission due to a deficit of high-quality experimental data.  Lastly, while all modeling of (n,x) reactions requires a good understanding of unbound nuclear states populated in (n,x) reactions, it remains one of the least understood topics in LENS.  In this section we explore all three of these topics these “crosscutting” topics and present plans to improve them.

@@

For fission theory, we can make a distinction in practice between description of the fission process up to scission (pre-scission physics), and the events that follow scission (post-scission physics). Eventually, one could envision a comprehensive model that accurately follows the entire fission process from the formation of the parent nucleus to the last decays of the final product nuclei within a single internally consistent theoretical framework. Such a model is not yet feasible so different physics models are used to treat pre- and post-scission processes. A wide range of theoretical approaches have been developed since the late 1930's to describe pre-scission physics, including scission-point models \cite{Wil76,Lem15}, microscopic-macroscopic models \cite{Moller2015} with random-walk dynamics \cite{Wa17}, Langevin dynamics \cite{Sie17}, Time-dependent Hartree-Fock and Hartree-Fock-Bogoliubov theory \cite{Koo76,Neg78,Tan15,God16}, the time-dependent superfluid local density approximation \cite{Ste11,Bul13,Ste15,Bul16}, and the time-dependent generator-coordinate method \cite{Ber91,Gou05,Rei16}.Several models have also been developed to describe post-scission physics and built into computer codes, such as GEF \cite{Sch16}, FIFRELIN \cite{Lit15}, CGMF \cite{Kaw10}, and FREYA \cite{Ran09,Ver18}.

@@

In addition to standard experiments looking at individual fission properties, multi-parameter measurements that give a comprehensive view of a fission event should be pursued as facilities and detectors improve. Ideally, a multi-parameter experiment would record all possible fission observables (e.g., fragment charge, mass, and energy distributions as well as neutron and gamma spectra and multiplicities) on an event-by-event basis, thereby making it possible to reconstruct the state of the nucleus at scission. In practice, each additional observable reduces statistics, making such measurements exceedingly challenging. Finally, there are experimental measurements that directly probe the pre-scission dynamics of the fission process although they carry large uncertainties and suffer from various ambiguities and limitations. These include measurements of scission neutrons \cite{Pe09}, fission time scales \cite{Jac09}, and muon-induced fission \cite{Mar80}. Probes of pre-scission physics can provide unique insights into the details of the fission process, and are therefore extremely valuable to fission theory.

@@

%Heading 1
\section{FIRST-LEVEL HEADING}
This is dummy text. 
% Heading 2
\subsection{Second-Level Heading}
This is dummy text. This is dummy text. This is dummy text. This is dummy text.

% Heading 3
\subsubsection{Third-Level Heading}
This is dummy text. This is dummy text. This is dummy text. This is dummy text.

% Heading 4
\paragraph{Fourth-Level Heading} Fourth-level headings are placed as part of the paragraph.

>>>>

\usepackage[american]{babel}
% \usepackage{amsmath}
\usepackage[version=3]{mhchem} 
% % \usepackage{fixltx2e}
% % \usepackage{refcount}
% % \usepackage{siunitx}
% % \usepackage{lastpage}
% % \usepackage{textcomp}
% \usepackage{mathtools}
% 
% \usepackage{xfrac}
\usepackage{lmodern}
\usepackage[hidelinks]{hyperref}
% % \usepackage{cool}
% % \usepackage{cancel}
\usepackage{microtype}
% \usepackage{listings}
% % \usepackage{mcode}
\usepackage [autostyle, english = american]{csquotes}
% % \usepackage{longtable}
% % \usepackage{subcaption}
% \usepackage{booktabs,siunitx}
% \usepackage{gensymb}
% \usepackage[normalem]{ulem}

@@

% \makeatletter
%  \def\@textbottom{\vskip \z@ \@plus 1pt}
%  \let\@texttop\relax
% \makeatother

@@

%Abstract
\begin{abstract}
A well-established knowledge of nuclear phenomena including fission, reaction cross sections, and structure/decay properties, is critical for applications ranging from the design of new reactors, to non-proliferation, to the production of radioisotopes for the diagnosis and treatment of illness.  However, the lack of a well-quantified, predictive theoretical capability means that most nuclear observables must be measured directly in order to \enquote{calibrate} empirical models, which, in turn, provide the data needed for these applications.  In many cases there is either a lack of the data needed to guide the models, or the results of the different measurements are discrepant.  This has led to the development of evaluation methodologies to provide recommended values and uncertainties.  In this article we present the nuclear data evaluation process and the international community that carries it out.  We then discuss new measurements and improved theory and/or modeling needed to address future challenges in applied nuclear science. 
\end{abstract}

@@

The nuclear structure compilation database is called the Experimental Unevaluated Nuclear Data List (XUNDL) \cite{XUN}.   The nuclear data in XUNDL is organized by nuclide, meaning that a single publication could lead to the production of more than one XUNDL dataset.  Each XUNDL dataset is treated as a \enquote{stand-alone} work and is not required to agree with existing, evaluated nuclear data for the nuclide it concerns.  However, during XUNDL compilation basic consistency checks are performed and any internal inconsistencies are pointed out to the author to allow them be reconciled in the database.  XUNDL is used by both nuclear structure evaluators to aid in their work and by nuclear structure researchers as a quick way to get access to data from current publications in order to guide their own research.

@@

While ENSDF and ENDF contain a vast quantity of nuclear data, there are significant amounts of nuclear data that are either not contained in them because they \enquote{fall between} nuclear structure and nuclear reactions, or they are present in a format that is not well-suited to their use in applications.  Examples of this include the recently updated Atlas of Neutron Resonances \cite{Mug18} which contains neutron capture resonance widths, centroids and cross sections and the Evaluated Gamma Activation File of (EGAF) capture gamma-rays \cite{Fir15} and the Atlas of Inelastic Scattering of Reactor Fast Neutron \cite{Hur18}.  Another very important nuclear data resource for applications is the Reference Input Parameter Library (RIPL) \cite{Capote2009} which contains both discrete (distilled from ENSDF) and continuous (Nuclear Level Densities and Gamma-strength function) nuclear structure data needed for accurate reaction modeling.

@@

The final step in the Pipeline is validation.  In this step, we use user codes and user-defined problems to validate that the data performs as expected.  The class of user problems we most rely on are so-called \enquote{integral benchmark} data that depend on many different types of data including cross sections and emitted particle spectra and are also extremely precisely known.  The canonical example of such a benchmark is a critical assembly, where the ratio of neutron production to neutron loss ($k_{eff}$) is known to several significant figures.  However, there are many other kinds of integral benchmarks used to test neutron interaction data alone, including reaction rate measurements and 14 MeV and fission spectrum transmission experiments, and many more for other forms of nuclear data.

There are multiple ways to interact with the nuclear data evaluation process along the way, including through web interfaces hosted by the National Nuclear Data Center \cite{NNDC} and the IAEA Nuclear Data Section \cite{IAEA-NDS}, a smartphone-based application \cite{IAEA-APP} and publication in peer-reviewed journals including Nuclear Data Sheets \cite{NDS} and Atomic Data and Nuclear Data Tables \cite{ATNDT}.

@@

Much of the work of nuclear data evaluation takes place as a part of international collaborations covered under the Organization of Economic Cooperation and Development's Nuclear Energy Agency (NEA) and the International Atomic Energy Agency (IAEA).  The Nuclear Energy Agency (NEA) addresses the nuclear technology interests of its 33 member states and, in the area of nuclear data, the Working Party on Evaluation Coordination (WPEC) is the main forum for collaborative effort between the nuclear data library projects from the NEA countries, namely ENDF/B (United States), JENDL (Japan), JEFF (NEA), TENDL (Europe), BROND (Russia) as well as the non-OECD file project CENDL (China).

A powerful instrument of WPEC to drive progress in nuclear data is the so-called Subgroup (SG): members of WPEC identify a common area in nuclear data that requires improvement, and if enough support from the various data library projects is present, a Subgroup is formed.  Subgroups typically operate on a 3-5 year time frame. Recent successful examples of WPEC subgroups, with specific relevance to the US, are large-scale horizontal efforts like CIELO (SG40), on worldwide nuclear data evaluation for the most important fission energy related materials. The CIELO initiative has, for example, led to new ENDF/B-VIII.0 evaluations for \ce{^{235,238}U}, \ce{^{56}Fe} and \ce{^{16}O} among others, and the Generalized Nuclear Database Structure (GNDS) format (SG38) \cite{SG38}, or more specific efforts like covariance adjustment for improvement of nuclear data files (SG39) \cite{SG39}.

WPEC also hosts long-term Expert Groups, and current groups are working on the development of the Generalized Nuclear Database Structure (GNDS) format, to provide a data library interface between nuclear physics and applications more modern than the ENDF-6 format, and the High-Priority Request List which assembles the most important nuclear data requests from applications in a unified format, to stimulate experimentalists and evaluators to provide these data. A full list of past and current subgroups of WPEC is available at Ref. \cite{WPEC}.

The International Atomic Energy Agency (IAEA) covers the interests of its 170 member states. The main task of the IAEA Nuclear Data Section is to provide fundamental nuclear data bases for basic and applied use, with data originating from experiments and theoretical simulations and covering both nuclear structure and nuclear reaction data.  An important collaboration coordinated by the IAEA is the Nuclear Reaction Data Center Network, which is responsible for keeping the EXFOR database of experimental nuclear reaction data up to date. NNDC is responsible for the US input to the database.

In addition, the IAEA organizes Coordinated Research Projects (CRP) and technical meetings as instruments to align international nuclear data efforts towards the production of validated databases ready for applied use.  Examples of recent and current CRP's include a 2018 venture centered on nuclear data for primary radiation damage has been completed, an ongoing effort to improve nuclear model parameters for fission reaction calculations by modern nuclear model codes such as EMPIRE \cite{Herman2007}, CCONE \cite{Iwa16}, COH3 \cite{KAWANO2010} and TALYS \cite{Koning2012} is ongoing, and an effort to create a first-ever evaluated database of radiative strength.  In 2019, a CRP on fission yields will start which aims to produce updated fission yield libraries for the major actinides to respond to requests from reactor technology, safeguards and non-proliferation.

@@

Accurate, physics-based modeling of neutron-induced reactions, particularly on the \enquote{Big 3} (\ce{^{235,238}U} and \ce{^{239}Pu}) are key to virtually all nuclear science and technology applications.  While a great deal of effort has been made to improve (n,f) data, a fundamental understanding of this most important of reaction mechanisms remain elusive.  Furthermore, since fission leads to the production of neutrons with energies up to several MeV, elastic and inelastic scattering also play a critical role.  Unfortunately, efforts to improve neutron scattering data has tended to \enquote{take a back seat} to fission due to a deficit of high-quality experimental data.  Lastly, while all modeling of (n,x) reactions requires a good understanding of unbound nuclear states populated in (n,x) reactions, it remains one of the least understood topics in LENS.  In this section we explore all three of these  \enquote{crosscutting} topics and present plans to improve them.

@@

For fission theory, we can make a distinction in practice between description of the fission process up to scission (pre-scission physics), and the events that follow scission (post-scission physics). Eventually, one could envision a comprehensive model that accurately follows the entire fission process from the formation of the parent nucleus to the last decays of the final product nuclei within a single internally consistent theoretical framework. Such a model is not yet feasible so different physics models are used to treat pre- and post-scission processes. A wide range of theoretical approaches have been developed since the late 1930's to describe pre-scission physics, including scission-point models \cite{Wil76,Lem15}, microscopic-macroscopic models \cite{Mol15} with random-walk dynamics \cite{Wa17}, Langevin dynamics \cite{Sie17}, Time-dependent Hartree-Fock and Hartree-Fock-Bogoliubov theory \cite{Koo76,Neg78,Tan15,God16}, the time-dependent superfluid local density approximation \cite{Ste11,Bul13,Ste15,Bul16}, and the time-dependent generator-coordinate method \cite{Ber91,Gou05,Rei16}.Several models have also been developed to describe post-scission physics and built into computer codes, such as GEF \cite{Sch16}, FIFRELIN \cite{Lit15}, CGMF \cite{Tal14}, and FREYA \cite{Ran09,Ver18}.

@@

In addition to standard experiments looking at individual fission properties, multi-parameter measurements that give a comprehensive view of a fission event should be pursued as facilities and detectors improve. Ideally, a multi-parameter experiment would record all possible fission observables (e.g., fragment charge, mass, and energy distributions as well as neutron and gamma spectra and multiplicities) on an event-by-event basis, thereby making it possible to reconstruct the state of the nucleus at scission. In practice, each additional observable reduces statistics, making such measurements exceedingly challenging. Finally, there are experimental measurements that directly probe the pre-scission dynamics of the fission process, although they carry large uncertainties and suffer from various ambiguities and limitations. These include measurements of scission neutrons \cite{Pe09}, fission time scales \cite{Jac09}, and muon-induced fission \cite{Mar80}. Probes of pre-scission physics can provide unique insights into the details of the fission process, and are therefore extremely valuable to fission theory.

@@

\subsection{Improved \ce{^{235,238}U} and \ce{^{239}Pu}(n,n') data for neutron transport}

Elastic and inelastic scattering have a profound effect on neutron populations in any nuclear application by changing the directions and energies of the neutrons in the system.  However, improvements in neutron scattering data have hampered by the fact that they do not produce easily measured signals like fission, or a radiochemical observable like neutron capture.  The result has been significant disagreements between different reaction evaluations.  These disagreements manifest themselves as compensating errors that are resistant to the validation process.  This issue has been recognized by the nuclear data community since the early days of the WPEC \cite{Row89} and more recently by Maslov \cite{Mas11}.

One compelling example of compensating errors involves the Jezebel 239Pu critical assembly (PMF1).  Models of Jezebel using LANL CIELO-1 (ENDF/B-VIII.0) and CEA CIELO-2 (JEFF-3.3) reproduce a unity value of keff to better than 1 part in 104.  However, when the authors substituted one nuclear quantity one quantity at a time between the two databases, the value of keff varied considerably.  The greatest changes came from changing the elastic and inelastic scattering cross sections, which led to changes in the keff of -500 pcm and +704 pcm respectively. This is shown in Fig. 2 to the right.   Similar examples of such compensating errors were also seen in earlier work by Bauge \cite{Bau12}.

A new Collaborative International Evaluated Library Organization (CIELO) Coordinated Research Program was started under the auspices of the IAEA \cite{Cha18a} to have the international nuclear data evaluation community agree on a set of evaluated cross sections and output spectra for six critically-important nuclei: 1H, 16O, 56Fe, 235,238U and 239Pu in order to address the issue of compensating errors.  Recently, this list has been expanded to include 14,15N, 9Be, 23Na, 59Co, 58Ni, 238-242Pu.

CIELO has relatively little energy or angle differential data to draw upon to aid in its evaluation of (n,n’) on the “Big 3” Actinide, leading to “significant differences amongst the evaluations for 239Pu in the fast energy range, with ENDF/B-VII.1 and JENDL-4.0 lying significantly above the JEFF-3.1 evaluation.” \cite{Cha14}.  The 239Pu data referenced in their paper are all from indirect measurements made prior to 1970 \cite{Bat69, And61}.  The case is similar for 235U where they note that differences for En > 50 keV “have a large impact on the fast criticality (Godiva), leading to a 540 pcm swing in calculated criticality, as shown by Go Chiba \cite{Chi12}.”

One significant source of (n,xn) partial cross section data from 1997 to 2003 was the GEANIE collaboration (Germanium Array for Neutron-Induced Excitations) at the Los Alamos Neutron Science Center/Weapons Neutron Research (LANSCE/WNR) facility. While the primary goal of the GEANIE collaboration was the determination of the 239Pu(n,2n) cross section \cite{Ber02}, it resulted in the publication of more than two dozen papers over the course of more than 15 years.  Experiments were run on all three of the major Actinides: 235U \cite{You01}, 239Pu \cite{Ber02} and 238U \cite{Fot04}.  It should be noted that GEANIE did not include any neutron detectors, producing only angle-integrated partial -ray cross sections.

The need for new scattering data has motivated many new experiments.  This has included energy-differential/angle-integrated measurements of the 238U(n,n’) cross section using the GELINA spallation neutron source in Geel Belgium by Bacquias et al., \cite{Nem13} and quasi-angle-differential measurements using the Rensselaer Polytechnic Institute Linac by Daskalakis et al., \cite{Das14}.

The ideal way to improve (n,n’) data on the 235,238U and 239Pu would be to perform measurements of the scattered neutron in coincidence with transitions emitted by the excited nucleus.  In the case of the Actinides this will potentially require the observation of conversion electrons given that the lowest transitions in all actinides are almost entirely converted.  Unfortunately, neutron beams usually have quite low intensity, with a fluence in the fast (1-10 MeV) region between 103 and 107 n/s/cm2 at the target location.  Furthermore, many spallation neutron sources have a significant high energy spectral component that can cause inter-element scattering in a neutron scintillator array rendering precise determination of the scattered neutron energy hard to  perform. However, intense limited-energy neutron sources based on thick-target deuteron breakup \cite{Har18} and forward-fitting techniques developed for use at spallation neutron sources \cite{Kee18} off an opportunity to perform neutron-gamma coincident measurements in the fast fission energy spectrum.  These techniques could provide valuable data to improve nuclear data evaluations of inelastic scattering, even on targets where fission is an open channel.

\subsection{Improved treatment of continuum nuclear data properties}

Information regarding states in the unresolved continuum region above the particle separation energy is of central importance to accurately modeling the competition between particle and gamma-ray emission.  The most important of these are the nuclear level density (NLD) as a function of energy, spin and parity, and the radiative strength function (RSF), which describes the ability of excited nuclear matter to absorb or emit photons.  A number of recent studies have shown the importance of the RSF to modeling (n,) reactions for neutrons with energies between 1 and 100 keV, which corresponds to the spectrum at the site of neutron capture nucleosynthesis in stars \cite{Lar10, Lar15} and energy generation in fast reactors and nuclear security applications.

Over the past several years’ work led by the group at the University of Oslo has shown that there is a surprisingly large amount of M1 radiative strength believed to be a type of “scissors” collective mode throughout the Actinide mass region near the valley of stability including 238U, 239,240,242Pu and 232Th \cite{Lap16, Gut13, Gut13a, Gut14, Gut15}.

Other results from the Oslo group show that Actinide NLDs appears to follow a constant temperature (CT) form with a roughly Gaussian dependence on angular momentum that is often described via an energy-dependent width spin cut-off parameter () of the form:

		(1)

where U is roughly the Ex minus the pairing gap , a is the level density parameter, A is the nuclear mass and is the asymptotic level density value one would obtain in the absence of any shell effects.

There is a great deal uncertainty in the magnitude of the spin cut-off parameter.  This is a significant issue since a good knowledge of the J distribution is needed to accurately model the neutron vs. -ray emission probability in nuclear reactions.  This sensitivity was shown recently by Wiedeking et al., \cite{Wie16} who showed that neutron emission from excited states in 95Mo well-above the neutron separation energy populated via (d,p) was hindered for states with high angular momentum.

(n,n’) provides a useful tool for studying the spin cut-off parameter since the reactions proceed largely through the population of a compound nucleus with a relatively wide range of angular momentum. Figure 4 shows the ratio of the partial cross section for the population of the  level in 238U(n,n’) as a function of incident neutron energy calculated using the TALYS reaction code for two different choices of spin cut-off.  The solid line shows the case where a constant temperature level density is assumed and the spin cut-off parameter was varied from 75% to 125% of its “normal” value given by the equation (3) above.  The ratio decreases by nearly a factor of 2 with increasing energy.  In contrast, the dashed line shows the ratio of the  population of the  state when the LD and RSF from the recent Oslo measurements \cite{Gut13a} is divided by the results obtained using the default Gilbert and Cameron in TALYS.  In this case the ratio decreases by nearly a factor of 2.  Similar behavior is seen in the partial cross sections for other off-yrast levels as well.

Information regarding the J-dependence of the NLD can also be obtained through observation of the relative population of two low-lying states with very different spins via activation.  This method has been used in Hg and Au by Chakravarty \cite{Cha92} and more recently by Sudar in Hg and Pt \cite{Sud06}.  Figure 5 shows this for the case of 52mMn(Jπ=2+)/52gMn(Jπ=6+) ratio populated via 56Fe(p,n) measured via stacked target activation at the LBNL 88-Inch cyclotron as compared to TALYS calculations for different values of the spin cut-off parameter (left) and the 6 level density models available for use in TALYS on the right.  
It should be noted that both the (n,n’) and isomer-to-ground state ratio measurements could be performed as part of a program that supports improved neutron transport as described in Section II.B above, and a program of cross section measurements in support of medical isotope production described in sections IV below.  This reflects the interdependency between nuclear data over a wide range of applications.

\section{NATIONAL SECURITY AND NON-PROLIFERATION}

\subsection{Interpreting the reactor anti-neutrino spectrum}

One canonical example of a non-proliferation activity where improvements in multiple cross-cutting areas of nuclear data are needed is the interpretation of the anti-neutrino spectrum from the prompt decay of fission fragments in the core of a nuclear reactor. Nuclear reactors are prolific sources of electron antineutrinos, producing about 1021 antineutrinos per second for a typical power reactor.  These electron antineutrinos are produced by the beta-minus decay of the more than 800 neutron-rich fission fragments, which are the debris from the main source of energy generation in a reactor, the neutron induced fission of actinide nuclides.  These antineutrinos are also the only radiation escaping from the vessel of a safely operating nuclear reactor, making them ideally suited for reactor monitoring, even in the event that the actor operating the reactor refuses to cooperate with the international community.  Recent technological advancements in electron antineutrino detection has made another dream a reality, monitoring the performance of nuclear reactors and tracking 239Pu for non-proliferation treaty verification.   The WATCHMAN experiment in the UK has just started to explore this field \cite{Cha18}.  Furthermore, the possible use of antineutrino detectors for the detection of nuclear explosions has been discussed \cite{CarAx}.

Nuclear reactors have also been an essential tool to study the weak interaction.  Their large antineutrino flux was capitalized by Cowan and Reines to discover antineutrinos in 1956 \cite{Cow56}, more than 25 years after they were first hypothesized by Pauli in 1930 to explain the continuum electron spectra observed following beta-minus decay.   In the last few years, the transformation of electron antineutrinos into the other two flavors was beautifully measured by three large-scale experimental efforts, Daya Bay \cite{An16}, Double Chooz \cite{Abe12} and RENO \cite{Cho16}.   These experiments also confirmed a deficit of antineutrinos of about 5-9% at short distances that had been revealed in a 2011 re-analysis of the conversion procedure to obtain antineutrino spectra from the measured electron spectra \cite{Men11}.

This intriguing deficit, as well as a spectrum distortion, has triggered a new generation of very-short distance reactor experiments, such as NEOS (Korea) \cite{Ko17}, DANSS (Russia) \cite{AleAx}, STEREO (France) \cite{AllAx}, PROSPECT (USA) \cite{Ash16}, and SoLid (Belgium) \cite{Kal17}, whose first results are beginning to be made public.  There are no signs of slowing down in the field of nuclear reactor antineutrinos, as the largest experiment to date, JUNO designed to measure the neutrino mass hierarchy should be ready for data taking by 2021 \cite{SalAx}.  JUNO will feature a 20 Million-Ton detector at 80 km from A 27 GW nuclear power plant in China.

The antineutrino spectrum produced in a nuclear reactor is calculated as the sum of the spectra produced by each of the nuclear fuels, 235,238U and 239,241Pu, weighted by the respective fission fractions \cite{Vog81}.   The spectra for each fuel can be calculated using nuclear databases, in what is commonly known as the ‘summation method’.   Alternatively, for 235U and 239,241Pu the spectra can also be calculated by converting the aggregate electron spectra measured at ILL in the 1980s \cite{Fel82,Sch85,Hah89}.   The current best predictions for the 235U and 239,241Pu are those by Huber \cite{Hub11} using the conversion method, while for 238U is that by Mueller et al. \cite{Mue11} using a hybrid conversion-summation method.

The use of nuclear databases to calculate antineutrino spectra is considered less precise than the conversion method due to deficiencies in the underlying fission and decay data, such as beta intensities \cite{20}, beta shape factors due to first-forbidden transitions \cite{Hay14}, evaluated libraries issues \cite{Son15} and fission yields \cite{Son16}.   There have been considerable efforts in the last few years to improve this situation, including (i) experimental campaigns in Jyvaskyla \cite{Alg10}, ORNL \cite{Ras16} and ANL \cite{MccTBD} to obtain highly precise values of beta intensities, which have been incorporated in the ENDF/B-VIII.0 decay data sub-library; (ii) measurements of isomeric ratios in ILL, Jyvaskyla and ANL; (iii) measurements of fission yields in GELINA \cite{Viv00}, GSI \cite{Pel17} and LANL \cite{Duk16}; (iv) a recent IAEA Coordinated Research Project on beta-delayed neutron emitters \cite{DimXX} that would improve the derivation of cumulative fission yields from independent ones.   Also planned are measurements of beta shape factors as well as a new IAEA Coordinated Research Project to evaluate fission yields.
Another issue affecting the summation method is the lack of fission yield correlation matrices in the evaluated fission data libraries.   As a result, uncertainties cannot be fully computed.  There has been some progress in this area, for instance, the ongoing OECD-NEA subgroup 44 will incorporate this topic \cite{SobXX}.

The summation method has been recently applied in two different situations.   First, it has been shown that the Daya Bay fuel evolution data \cite{An17} can be explained by summation calculations including an anomaly of about 4% \cite{Hay18}.   Second, a novel method to analyze the Daya Bay spectrum revealed the signature of 4 individual fission products, while the same method applied to the electron spectra revealed two additional fission products \cite{Son18}.

The conversion method is not fully free of issues, either.  To start, it is based on a single set of measurements; also, its most basic premise, the description of a level to level spectrum using a particular formalism has not been tested as highly precise measurements of electron spectra for fission products above the inverse beta decay cross section threshold have not been published.    As a result, while more precise than the summation method, it is nevertheless considerably less reliable.     A recent article \cite{Son17} reviewed the different Fermi functions available for the calculation of spectra and showed that a linear correction factor of 6%/MeV could explain the anomaly.

Given the dependence of the antineutrino signal on both decay and fission product yield data a series of sensitivity study-guided integral experiments, along the lines of critical assembly integral benchmarks, to validate the underlying nuclear data would provide a strong constraint on the underlying data.  One approach would involve irradiating fissile and fissionable samples in facilities with well-understood neutron spectra and comparing the observed delayed gamma-rays to a modeled spectrum produced using the Fission Induced Electromagnetic Response (FIER) code developed at Berkeley \cite{Mat18}.  FIER convolves fission product yield distributions and decay data to predict the gamma-ray spectrum after neutron irradiation.  The first example of its use involving a comparison to the delayed gamma-ray spectrum from a 235U foil irradiated in the Godiva critical assembly indicated several cases were the underlying fission yields or decay data required improvement.

These experiments would involve a sample of depleted 238U since recent results from IBD indicate that the contribution to the observed anomaly in the antineutrino experiment can be attributed to issues with the 238U(n,f) yields. These measurements could be performed at the Godiva in the same fashion as the 235U irradiation used to validate FIER, or other well-documented fast neutron facilities including the thick target deuteron break neutron source at the LBNL 88-Inch cyclotron \cite{Har18} and the High Flux Neutron Generator at UC-Berkeley \cite{All18} and compare the results to the FIER predictions to inform both the fission yield and decay data for 238U(n,f).

\subsection{Improved (n,x) data for active interrogation}

An invaluable tool for determining the presence of fissile materials for non-proliferation involves prompt neutron- and gamma-ray spectroscopy coupled to a DT neutron sources with associated particle imaging (DT-API).   DT-API sources allow tracking of the trajectory of individual neutrons, thereby greatly decreasing background from scattered neutrons.  The signal from DT-API generators includes long-chain gamma-ray emission from multiple fission events \cite{Pra12,Nak10} as well as prompt gamma- and neutron scattering on a wide range of low- and high-Z materials.

The interpretation of this data requires detailed knowledge of both the double-differential neutron scattering and gamma-ray production cross sections from fast down to thermal neutron energies.  The Evaluated Gamma Activation File \cite{Fir14} provides partial gamma-ray cross section data for transitions between discrete states following thermal neutron capture.  At higher energies the Atlas of Gamma-rays from the Scattering of Reactor Fast Neutrons \cite{Dem78} which has recently been compiled into a database and reconciled to reflect state-of-the-art gamma-ray energies from ENSDF \cite{Hur18} offers discrete gamma-ray cross sections for irradiation with fast fission neutrons.

A program of measurements of gammas emitted from both discrete and QC states would require a relatively modest effort to set in place.  The data from these measurements could then be used to guide physics-based reaction model, such as the COH, TALYS or EMPIRE to ensure that the total gamma-ray production cross section is consistent with the total capture and scattering cross sections themselves.

\section{ISOTOPE PRODUCTION}
